---
title: "SVM Example"
author: "Will Doyle"
date: "11/15/2022"
output: html_document
---
```{r}
library(tidyverse)
library(tidymodels)
library(finetune)
library(janitor)
```

## Read in Data
```{r}
cars<-read_csv("../data/training_car.csv")%>%
  clean_names()%>%
  select(is_bad_buy,
         veh_year,
         make,
         color,
         transmission,
         veh_odo,
         is_online_sale)%>%
         mutate(is_bad_buy=as_factor(
           recode(is_bad_buy,
                           `1`="BadBuy",
                           `0`="GoodBuy")))

cars<-cars%>%sample_n(1000)  
```

## Split Train and Test

```{r}
cars_split<-initial_split(cars)

cars_train<-training(cars_split)

cars_test<-testing(cars_split)
```



## Set up for resampling 

The following code sets up for monte carlo resampling, with 100 reps

```{r}
cars_rs<-mc_cv(cars_train)
```

## Set Formula
```{r}
car_formula<-as.formula("is_bad_buy~.")
```

## Set recipe
```{r}
cars_rec<-recipe(car_formula,cars_train)%>%
  update_role(is_bad_buy,new_role = "outcome")%>%
  step_naomit(all_predictors())%>%
  step_other(all_nominal_predictors(),threshold = .05)%>%
  step_dummy(all_nominal_predictors())%>%
  step_normalize(all_predictors())
```

## SVM Specifications

Below, I set up two svm specifications, one with a polynomial kernal (svm_poly) and one with a radial basis function kernel (rbf). 

```{r}
## Polynomial kernel
svm_poly_spec <-
  svm_poly(degree = tune(),scale_factor=tune(),cost=tune()) %>%
  set_mode("classification") %>%
  set_engine("kernlab")

## Radial basis function kernel
svm_rbf_spec <-
  svm_rbf(rbf_sigma = tune(),cost=tune())%>%
  set_mode("classification") %>%
  set_engine("kernlab")
```


## Set workflows
```{r}
poly_wf<-workflow()%>%
  add_model(svm_poly_spec)%>%
  add_recipe(cars_rec)
```

```{r}
rbf_wf<-workflow()%>%
  add_model(svm_rbf_spec)%>%
  add_recipe(cars_rec)
```

## Simulated Annealing

Below, I use a different process for tuning-- simulated annealing. Instead of specifying the full grid, I'm allowing the tuning algorithm to walk through a random space in order to find the best possible combination of tuning parameters. 

https://www.youtube.com/watch?v=qEeF-ErtUAU&t=99s

```{r}
set.seed(123)
doParallel::registerDoParallel()

cars_poly_results <-
  poly_wf%>%
  tune_sim_anneal(
    resamples=cars_rs,
    iter=50
  )
```


Same thing, with radial basis function kernel. 
```{r}
set.seed(123)
doParallel::registerDoParallel()

cars_rbf_results <-
    rbf_wf%>%
  tune_sim_anneal(
    resamples=cars_rs,
    iter=50
  )
```


Here's the results for polynomial kernel.
```{r}
cars_poly_results%>%
  collect_metrics()%>%
  filter(.metric=="roc_auc")%>%
  arrange(-mean)
```

And for the radial basis function kernel.
```{r}
cars_rbf_results%>%
  collect_metrics()%>%
  filter(.metric=="roc_auc")%>%
  arrange(-mean)
```

We can also take a look via `show_best`
```{r}
show_best(cars_rbf_results,n = 20,metric="roc_auc")
```




